---
title: "p8105_hw2_hz2710"
author: "Huili Zheng"
date: "10/7/2021"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(readxl)
library(dplyr)
library(lubridate)
library(ggplot2)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```


## Problem 1


Read and clean the Mr. Trash Wheel sheet:

```{r}
wheel_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",range = "A2:N406" ) %>%
  janitor::clean_names() %>%
  drop_na("dumpster") %>%
  mutate(
    sports_balls = round(sports_balls)
  ) 
```


Read and clean precipitation data for 2018 and 2019. 

```{r}
pre2018_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2018 Precipitation",
             range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na("total") %>%
  mutate(
    year = 2018
  )
pre2019_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2019 Precipitation",
             range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na("total") %>%
  mutate(
    year = 2019
  )

pre_df = 
  left_join(pre2018_df, pre2019_df, by = "month") %>%
  mutate(
    mouth = month.name[month]
  )
```


The number of observations in Mr.Trash Wheel dataset is `r nrow(wheel_df)` and the number of observations in precipitation data is `r nrow(pre_df)`. The key variables of Mr.Trash Wheel dataset include "weight_tons" and "volume_cubic_yards". The key variables of precipitation data are "month" and "total". The total precipitation in 2018 is `r sum(pull(pre2018_df,total))`. The median number of sports balls in a dumpster in 2019 is `r median(pull(filter(wheel_df,year == 2019), sports_balls))`.


## Problem 2
This problem uses the FiveThirtyEight data; these data were gathered to create the interactive graphic on this page. In particular, we’ll use the data in pols-month.csv, unemployment.csv, and snp.csv. Our goal is to merge these into a single data frame using year and month as keys across datasets.

First, clean the data in pols-month.csv. 

```{r}
pols_month_df = 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c('year', 'month', 'day'), sep = "-") %>%
  mutate(
    year = as.integer(year),
    month = month.name[as.integer(month)],
    day = as.integer(day),
    president = case_when(prez_gop == 1 ~ "gop" ,
                          prez_dem == 1 ~ "dem"),
  ) %>%
  select(-c(day, prez_dem, prez_gop))


```

Second, clean the data in snp.csv using a similar process to the above. 

```{r}
snp_df = 
  read_csv("data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  mutate(date = mdy(date)) %>%
  separate(date, into = c('year', 'month', 'day'),sep = "-") %>%
  mutate(
    year = as.integer(year),
    month = month.name[as.integer(month)],
  ) %>%
  relocate(year, month) %>%
  select(-(day))
```

Third, tidy the unemployment data so that it can be merged with the previous datasets. 

```{r}
unemployment_df = 
  read.csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(jan:dec, names_to = "month", values_to = "unemployment rate") %>%
  mutate(
    month = case_when(month == "jan" ~ "1",
                      month == "feb" ~ "2",
                      month == "mar" ~ "3",
                      month == "apr" ~ "4",
                      month == "may" ~ "5",
                      month == "jun" ~ "6",
                      month == "jul" ~ "7",
                      month == "aug" ~ "8",
                      month == "sep" ~ "9",
                      month == "oct" ~ "10",
                      month == "nov" ~ "11",
                      month == "dec" ~ "12"),
    month = month.name[as.integer(month)]
  )
```

Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
result_df = left_join(left_join(snp_df,pols_month_df,by = c("year","month")), unemployment_df, by = c("month","year"))

```

The pols-month dataset contains the number of governors, senators, representatives from republican and democratic parties from 1947 to 2015 and specifies the party the presidents were on the associated date.
The snp dataset contains the closing value of the S&P stock index on the associated date. 
The unemployment dataset contains the every month unemployment rate in the associated year.
The dimension of result dataset is `r nrow(result_df)` rows and `r ncol(result_df)` columns. The range of years is `r min(pull(result_df, year)) - max(pull(result_df, year))`, the names of key variables are "year" and "month".

## Problem3
This problem uses data from NYC Open data on the popularity of baby names.

Load and tidy the data. Address the problems of the names of a categorical predictor and the case structure of string variables.
```{r}
pop_name = 
  read_csv("data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(
    childs_first_name = str_to_lower(childs_first_name),
    gender = str_to_lower(gender),
    ethnicity = str_to_lower(ethnicity)
  ) %>%
  distinct()

count(pop_name, ethnicity)
pop_name = pop_name %>%
  mutate(
    childs_first_name = str_to_title(childs_first_name),
    ethnicity = case_when(
      ethnicity == "asian and paci" ~ "asian and pacific islander",
      ethnicity == "black non hisp" ~ "black non hispanic",
      ethnicity == "white non hisp" ~ "white non hispanic",
      ethnicity == "asian and pacific islander" ~ "asian and pacific islander",
      ethnicity == "black non hispanic" ~ "black non hispanic",
      ethnicity == "white non hispanic" ~ "white non hispanic",
      ethnicity == "hispanic" ~ "hispanic")
  )
```

Produce a table showing the rank in popularity of the name “Olivia” as a female baby name over time.
```{r}
olivia_df = 
  pop_name %>%
  filter(childs_first_name == "Olivia", gender == "female") %>%
  select(year_of_birth, ethnicity,rank) %>%
  pivot_wider(names_from = year_of_birth, values_from = rank) %>%
  knitr::kable()

olivia_df
```

Investigate the most popular name over time among male babies.
```{r}
## find the most popular male name
male_name_df = 
  pop_name %>% 
  filter(gender == "male") %>%
  group_by(childs_first_name) %>%
  summarise(frequency = sum(count)) %>%
  arrange(desc(frequency))

## the most popular name among male is ethan.
ethan_df = 
  pop_name %>%
  filter(childs_first_name == "Ethan", gender == "male") %>%
  select(year_of_birth, ethnicity,rank) %>%
  pivot_wider(names_from = year_of_birth, values_from = rank) %>%
  knitr::kable()

ethan_df
```

For male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
M_W_2016 = 
  pop_name %>%
  filter(gender == "male",  year_of_birth == 2016, ethnicity == "white non hispanic")
gg_rank_count = 
  M_W_2016 %>%
  ggplot(aes(x = rank, y = count)) +
  geom_point() +
  labs(
    title = "Baby names popularity",
    x = "Baby names rank",
    y = "Baby names count"
  ) 

gg_rank_count

ggsave("M_W_2016.jpg",gg_rank_count,device = "jpg")
```


