---
title: "p8105_hw2_hz2710"
author: "Huili Zheng"
date: "10/7/2021"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(readxl)
```


## Problem 1


Read and clean the Mr. Trash Wheel sheet:

```{r}
wheel_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",range = "A2:N406" ) %>%
  janitor::clean_names() %>%
  drop_na("dumpster") %>%
  mutate(
    sports_balls = round(sports_balls)
  ) 
```


Read and clean precipitation data for 2018 and 2019. 

```{r}
pre2018_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2018 Precipitation",
             range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na("total") %>%
  mutate(
    year = 2018
  )
pre2019_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2019 Precipitation",
             range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na("total") %>%
  mutate(
    year = 2019
  )

pre_df = 
  left_join(pre2018_df, pre2019_df, by = "month") %>%
  mutate(
    mouth = month.name[month]
  )
```


The number of observations in Mr.Trash Wheel dataset is `r nrow(wheel_df)` and the number of observations in precipitation data is `r nrow(pre_df)`. The key variables of Mr.Trash Wheel dataset include "weight_tons" and "volume_cubic_yards". The key variables of precipitation data are "month" and "total". The total precipitation in 2018 is `r sum(pull(pre2018_df,total))`. The median number of sports balls in a dumpster in 2019 is `r median(pull(filter(wheel_df,year == 2019), sports_balls))`.
